Fluent statistical computing interfaces for biological data analysis

### 2020-08-04 Rmd/00-e-preface.Rmd
style

This thesis has been written using R Markdown with the **bookdown**
package [@Xie2016-oa; @Xie2017-ke]. The **renv** package
has been used to create a reproducible environment to build the thesis from source (available at https://github.com/sa-lee/thesis ) [@renv]. An online version of the thesis is available to read at https://sa-lee.github.io/thesis.

>>>>

This thesis has been written using R Markdown with the **bookdown**
package [@Xie2016-oa; @Xie2017-ke]. The **renv** package
has been used to create a reproducible environment to build the thesis from source [@renv]. All materials required to compile the thesis are available at https://github.com/sa-lee/thesis. An online version of the thesis is available to read at https://sa-lee.github.io/thesis.

### 2020-08-04 Rmd/01-intro.Rmd
style

Chapter \@ref(ch:tsne) introduces a software package called **liminal** for 
constructing these views and a user interaction framework for identifying 
distortions. We present several case studies using data sets  both 
single-cell transcriptomics and  machine learning  for using our approach to diagnose the quality   results obtained via popular NLDR methods
like t-distributed stochastic neighbour embeddings (t-SNE) [@Maaten2008-sk].

>>>>

Chapter \@ref(ch:tsne) introduces a software package called **liminal** for 
constructing these views and a user interaction framework for identifying 
distortions. We present several case studies using data that capture aspects of
single cell transcriptomics workflows, and use our approach to diagnose the
quality of results obtained  via popular NLDR methods like t-distributed 
stochastic neighbour embeddings (t-SNE) [@Maaten2008-sk].

### 2020-08-04 Rmd/02-plyranges.Rmd
style

We have created a genomic DSL called **plyranges** that reformulates
notions from existing genomic algebras and embeds them in R as a
genomic extension of **dplyr**. By analogy, **plyranges** is to the
genomic algebra, as **dplyr**` is to the relational algebra.  The
**plyranges`**Bioconductor package implements the language on top of a
key subset of Bioconductor data structures and thus fully integrates
with the Bioconductor framework, gaining access to its scalable data
representations and sophisticated statistical methods.

@@

To support merging, our algebra recasts finding overlaps or 
nearest neighbors between two
genomic regions as variants of the relational join operator. A join
acts on two _GRanges_ objects: _x_ and _y_. The join operator
is relational in the sense that metadata from the _x_ and _y_
ranges are retained in the joined range.  All join operators in the
`plyranges` DSL generate a set of hits based on overlap or proximity
of ranges and use those hits to merge the two datasets in different
ways. There are four supported matching algorithms: _overlap_,
_nearest_, _precede_, and _follow_ (figure 
\@ref(fig:olaps-fig)). We can further restrict the
matching by whether the query is completely _within_ the subject, and
adding the _directed_ suffix ensures that matching ranges have the
same direction (strand).

@@

A caveat to constructing a compatible interface with **dplyr** is that
**plyranges** makes extensive use of non-standard evaluation in R via
the **rlang** package [@R-rlang]. Simply, this means that computations are
evaluated in the context of the _GRanges_ objects. Both **dplyr** and
**plyranges** are based on the **rlang** language, because it allows for more
expressive code that is free of repeated references to the
container. Implicitly referencing the container is particularly
convenient when programming interactively.  Consequently, when
programming with **plyranges**, a user needs to generally understand
the **rlang** language and how to adapt their code accordingly. Users
familiar with the tidyverse should already have such knowledge.

>>>>

We have created a genomic DSL called **plyranges** that reformulates
notions from existing genomic algebras and embeds them in R as a
genomic extension of **dplyr**. By analogy, **plyranges** is to the
genomic algebra, as **dplyr** is to the relational algebra.  The
**plyranges** Bioconductor package implements the language on top of a
key subset of Bioconductor data structures and thus fully integrates
with the Bioconductor framework, gaining access to its scalable data
representations and sophisticated statistical methods.

@@

To support merging, our algebra recasts finding overlaps or 
nearest neighbors between two
genomic regions as variants of the relational join operator. A join
acts on two _GRanges_ objects: _x_ and _y_. The join operator
is relational in the sense that metadata from the _x_ and _y_
ranges are retained in the joined range.  All join operators in the
**plyranges** DSL generate a set of hits based on overlap or proximity
of ranges and use those hits to merge the two datasets in different
ways. There are four supported matching algorithms: _overlap_,
_nearest_, _precede_, and _follow_ (figure 
\@ref(fig:olaps-fig)). We can further restrict the
matching by whether the query is completely _within_ the subject, and
adding the _directed_ suffix ensures that matching ranges have the
same direction (strand).

@@

A caveat to constructing a compatible interface with **dplyr** is that
**plyranges** makes extensive use of non-standard evaluation in R via
the **rlang** package [@R-rlang]. Simply, this means that computations are
evaluated in the context of the _GRanges_ objects. Both **dplyr** and
**plyranges** are based on the **rlang** language, because it allows for more
expressive code that is free of repeated references to the
container. Implicitly referencing the container is particularly
convenient when programming interactively.  Consequently, when
programming with **plyranges**, a user needs to generally understand
the **rlang** language and how to adapt their code accordingly. Users
familiar with the **tidyverse** should already have such knowledge.
